---
title: "N741pkg Vignette - Demonstration of the Functions"
author: "Melinda K Higgins"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Demo 1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Functions Included

To date, this `N741pkg` package includes 2 functions `tbl.continuous` and `tbl.cat` for creating summary statistics tables for numeric (continuous) data and categorical (or ordinal) variables.

### Creating summary statistics tables for numeric (continuous) variables

The `tbl.continuous` function will work for any numeric variable. Ideally, you will run these statistics on continuous measures but can compute these summary statistics for ordinal data as well. The resulting `data.frame` output from this function can be used easily with `knitr::kable()` to create a table of summary statistics. This summary includes both informative descriptives (sample size, missing data, minimum, maximum), parametric statistics (mean, standard deviation), and non-parametric statistics (median, 25th and 75th percentile for the interquartile range).

* `tbl.continuous(x,x$a,"label for var a")`
    - This function creates a table of summary statistics for a numeric variable in a data frame.
    - There are 3 arguments you need to provide:
        - a data.frame, `df`
        - a column variable from the data.frame `df$var`
        - a character vector to serve as a label for the variable `"var label"`
    - The returned table is a `data.frame` object with 1 row and 10 variables/columns. These 10 columns include:
        - `item`: the variable label provided
        - `n`: the number of non-missing values
        - `missing`: the number of missing values
        - `min`: the minimum value
        - `avg`: the average/mean value
        - `SD`: the standard deviation of the values
        - `median`: the median value
        - `Q1`: the 25th percentile (1st quartile) of the values
        - `Q3`: the 75th percentile (3rd quartile) of the values
        - `max`: the maximum value
        
### Example using the `mtcars` dataset with `tbl.continuous`

Using the `mtcars` built-in dataset, you can create separate `data.frame` object "tables" (t1, t2, ...) and then combine these using `rbind()` (bind data.frame objects together by rows) to make a bigger `data.frame` containing the summary stats for multiple variables at once which can be printed using `knitr::kable()`. Each row of the resulting table is a different variable (column) containing the summary statistics for that variable.

```{r}
m <- mtcars
t1 <- N741pkg::tbl.continuous(m,m$mpg,"Miles per Gallon")
t2 <- N741pkg::tbl.continuous(m,m$disp,"Engine Displacement")
t3 <- N741pkg::tbl.continuous(m,m$wt,"Car Weight")
t4 <- N741pkg::tbl.continuous(m,m$hp,"Horsepower")
t5 <- N741pkg::tbl.continuous(m,m$qsec,"1/4 mile time")
knitr::kable(rbind(t1,t2,t3,t4,t5),
             caption = "Table of Summary Stats for Numeric Variables in mtcars")
```

        
### Creating summary statistics tables for categorical (or ordinal) variables
        
* `tbl.cat(gx,gx$b)`
    - This function creates a table of summary statistics (frequencies and percents) for a categorical (or ordinal) variable in a data frame.
    - There are 2 arguments you need to provide:
        - a "grouped" data.frame, `gdf` **SEE NOTE**
        - a column variable from the "grouped" data.frame `gdf$var`
    - **NOTE** A "grouped" data.frame can be created using the `dplyr::group_by()` function. See the example provided with `help(tbl.cat)`. You need to "group_by" the categorical (or ordinal) variable you want to make the table for.
    - The returned table is a `data.frame` object with 1 row and 3 variables/columns. These 3 columns include:
        - `var`: the categorical variable categories or ordinal levels
        - `freq`: the frequencies (or counts) for each category or level
        - `pct`: the percentage of the total number of rows - these percents are NOT adjusted for missing. However, the frequency and percentage of `NA`s are provided.
        
### Example using the `mtcars` dataset with `tbl.cat`

In the `mtcars` dataset, there are several ordinal variables for which running frequency summaries is useful. Let's run frequency summaries for:

* `cyl` "Number of Cylinders"
* `am` "Transmission (0=automatic, 1=manual)"
* `gear` "Number of forward gears"

```{r}
# create grouped data.frame of m by cylinders
gm <- dplyr::group_by(m,cyl)
t1 <- N741pkg::tbl.cat(gm,gm$cyl)
knitr::kable(t1,
             caption = "Frequency Table for Number of Cylinders")

# create grouped data.frame of m by am
gm <- dplyr::group_by(m,am)
t1 <- N741pkg::tbl.cat(gm,gm$am)
knitr::kable(t1,
             caption = "Frequency Table for Transmission")

# create grouped data.frame of m by gear
gm <- dplyr::group_by(m,gear)
t1 <- N741pkg::tbl.cat(gm,gm$gear)
knitr::kable(t1,
             caption = "Frequency Table for Number of Forward Gears")
```


## Generic Vignette information - included for further examples

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
