---
title: "Vignette Demo the Functions in N741pkg"
author: "Melinda K Higgins"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Demo 1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

## test a continuous var table

```{r}
library(N741pkg)
a <- runif(10,0,1)
b <- c("a","b","c","d","a","a","b","b","c","c")
x <- data.frame(a,b)
c <- c("blue","blue","blue","blue","red",
       "green","green","green","black","orange")
d <- c(2,40,3,NA,5,1,2,4,3,NA)
x <- data.frame(a,b,c,d)

knitr::kable(x)

t <- N741pkg::tbl.continuous(x,x$d,"label ddd")
knitr::kable(t)
```


## test categorical table

```{r}
gx <- dplyr::group_by(x,c)
t <- N741pkg::tbl.cat(gx,gx$c)
knitr::kable(t)

gx <- dplyr::group_by(x,as.factor(d))
t <- N741pkg::tbl.cat(gx,gx$d)
knitr::kable(t)

```

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
